Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-04-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update information from Favela

\begin{sidewaystable}[]
\centering
\caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
$^a$ The number of lipid molecules
$^b$ The number of cholesterol molecules
$^c$ Cholesterol concentration (mol\%)
$^d$ The number of water molecules
$^e$ Simulation temperature
$^f$ The total simulation time
$^g$ Time frames used in the analysis
$^h$ Reference link for the downloadable simulation files
$^i$ Reference for the full simulation details
}\label{systemsCHOL}
\begin{tabular}{c c c c c c c c c c c}
%\hline
Force field & lipid   & $^a$N$_{\rm l}$ & $^b$N$_{\rm chol}$ &$^c$C$_{\rm CHOL}$  &  $^d$N$_{\rm w}$ & $^e$T (K)  & $^f$t$_{{\rm sim}}$(ns)  & $^g$t$_{{\rm anal}}$ (ns)& $^h$Files  &  $^i$Details\\
\hline
Berger-POPC-07~\cite{ollila07a}&   POPC &128 & 0 &0\% & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
/H\"oltje-CHOL-13~\cite{holtje01,ferreira13}   &    & &  &   &   &  &  &  &  \\
                               &   POPC &120 & 8 & 6\% &7290   & 298  & 100 & 80 & [\citenum{bergerFILESpopc7chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &110 & 18& 14\% & 8481  & 298  & 100 & 80 & [\citenum{bergerFILESpopc15chol}]$^*$ & [\citenum{ferreira13}]  \\
                               &   POPC &84 & 44 & 34\%  & 6794   & 298  & 100 & 80 & [\citenum{bergerFILESpopc34chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &64 & 64 & 50\% & 10314  & 298  & 100 & 80 & [\citenum{bergerFILESpopc50chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
CHARMM36\cite{klauda10,lim12}   & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  & SI  \\
                                & POPC   & 512& 0& 0\% & ?  & 298  & 170 & 100 & ?  & ?  \\
                                & POPC   & ?& ?& 10\% & ?  & 298  & 170 & 100 & ?  & ?  \\
                                & POPC   & ?& ?& 15\% & ?  & 298  & 170 & 100 & ?  & ?  \\
                                & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & SI \\
                                & POPC   & ?& ?& 20\% & ?  & 298  & 170 & 100 & ?  & ?  \\
                                & POPC   & ?& ?& 25\% & ?  & 298  & 170 & 100 & ?  & ?  \\
                                & POPC   & ?& ?& 35\% & ?  & 298  & 170 & 100 & ?  & ?  \\
                                & POPC   & 256& 256& 50\% & ?  & 298  & 170 & 100 & ?  & ?  \\
                               & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & SI \\
MacRog\cite{maciejewski14}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI \\ 
                          & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                          & POPC   & 72   & 56 &  44\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                             & POPC   & 64  & 64 & 50\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                             & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
\end{tabular}
\end{sidewaystable}

@@

{\it DPPC}
\todo{Markus Miettinen, please describe how you made the files.}

@@

{\it POPC}
The starting structures for the pure POPC and DOPC simulations was taken from the Slipids~\cite{jambeck12b} website (http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html).
The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
They contained 100 POPC/24 cholesterol molecules and 80 POPC/80 cholesterol molecules for
the simulations of 20\% cholesterol and 50\% cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
solvate the system.
The publicly available CHARMM36 forcefield parameters (\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
Single point energy calculation was done to assess the conversion. 
%(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
%http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
Simulations were performed for 200~ns and the last 100~ns was used for the calculations. Timestep of 2~fs was
used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
303~K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2~ps. Pressure was maintained semiisotropically
at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbour list with a cut-off of 1.2~nm
was updated every 10 steps. Lennard-Jones interactions were switched to zero
between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.

>>>>

\begin{sidewaystable}[]
\centering
\caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
$^a$ The number of lipid molecules
$^b$ The number of cholesterol molecules
$^c$ Cholesterol concentration (mol\%)
$^d$ The number of water molecules
$^e$ Simulation temperature
$^f$ The total simulation time
$^g$ Time frames used in the analysis
$^h$ Reference link for the downloadable simulation files
$^i$ Reference for the full simulation details
}\label{systemsCHOL}
\begin{tabular}{c c c c c c c c c c c}
%\hline
Force field & lipid   & $^a$N$_{\rm l}$ & $^b$N$_{\rm chol}$ &$^c$C$_{\rm CHOL}$  &  $^d$N$_{\rm w}$ & $^e$T (K)  & $^f$t$_{{\rm sim}}$(ns)  & $^g$t$_{{\rm anal}}$ (ns)& $^h$Files  &  $^i$Details\\
\hline
Berger-POPC-07~\cite{ollila07a}&   POPC &128 & 0 &0\% & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
/H\"oltje-CHOL-13~\cite{holtje01,ferreira13}   &    & &  &   &   &  &  &  &  \\
                               &   POPC &120 & 8 & 6\% &7290   & 298  & 100 & 80 & [\citenum{bergerFILESpopc7chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &110 & 18& 14\% & 8481  & 298  & 100 & 80 & [\citenum{bergerFILESpopc15chol}]$^*$ & [\citenum{ferreira13}]  \\
                               &   POPC &84 & 44 & 34\%  & 6794   & 298  & 100 & 80 & [\citenum{bergerFILESpopc34chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &64 & 64 & 50\% & 10314  & 298  & 100 & 80 & [\citenum{bergerFILESpopc50chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
CHARMM36\cite{klauda10,lim12}   & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  & SI  \\
                                & POPC   & 512& 0& 0\% & 23943  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 460& 52& 10\% & 23569  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 436& 76& 15\% & 23331  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & SI \\
                                & POPC   & 410& 102& 20\% & 20972  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 384& 128& 25\% & 22327  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 332& 180& 35\% & 21340  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 256& 256& 50\% & 20334  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                               & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & SI \\
MacRog\cite{maciejewski14}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI \\ 
                          & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                          & POPC   & 72   & 56 &  44\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                             & POPC   & 64  & 64 & 50\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                             & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
\end{tabular}
\end{sidewaystable}

@@

{\it DPPC and POPC with 72 lipids}
\todo{Markus Miettinen, please describe how you made the files.}

@@

{\it POPC with 128 lipids}
The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html).
The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
They contained 100 POPC/24 cholesterol molecules and 80 POPC/80 cholesterol molecules for
the simulations of 20\% cholesterol and 50\% cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
solvate the system.
The publicly available CHARMM36 forcefield parameters (\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
Single point energy calculation was done to assess the conversion. 
%(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
%http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
Simulations were performed for 200~ns and the last 100~ns was used for the calculations. Timestep of 2~fs was
used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
303~K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2~ps. Pressure was maintained semiisotropically
at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbour list with a cut-off of 1.2~nm
was updated every 10 steps. Lennard-Jones interactions were switched to zero
between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.

{\it POPC with 512 lipids}
POPC The starting structures for simulations were constructed with
the CHARMM-GUI website~\cite{jo08}. 512 POPC lipids (256 per leafleat) were 
used for the initial 0\% concentration and they were subsequently 
substituted by cholesterol molecules to fulfil the desired concentration.
The TIP3P~\cite{jorgensen83} water model was used to solvate the system. 
The publicly available port of the CHARMM36 forcefield parameters (\url{mackerell.umaryland.edu/CHARMM_ff_params.html}) 
was used for both POPC and cholesterol. 
Simulations were performed for 170 ns and the last 100 ns was used for the calculations.
Timestep of 1 fs was used with the leapfrog integrator. Covalent bonds with hydrogen 
were constrained with LINCS algorithm. Coordinates were written every 10 ps.
PME with real space cut-off of 1.2 nm was used for electrostatics. Lennard-Jones 
interactions were switched to zero between 0.8 nm and 1.2 nm. The neighbour lists with a
cut-off of 1.2 nm were updated every 5 steps. Temperature was coupled separately for cholesterol, lipids
and water to 298 K using the velocity-rescale method with coupling constant of 0.2 ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen
method.
